{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-06-15 thesis/education.tex
describe mysterious test case annos in landslide chapter

\section{P2}

@@

\subsection{Test cases}

Landslide ships with several ``approved'' test cases,
i.e., programs copied from, derived from, or at least vaguely resembling
the tests distributed with P2,
which I curated to produce concurrent behaviour suitable for stateless model checking.
Some tests are crafted to target specific bugs which,
from personal experience as a TA, are common in many student submissions;
others are crafted to exercise generally concurrency-heavy code paths and uncover any number of unforeseen problems.

@@

\subsubsection{Magic post-test assertions}

Test cases may define global variables of the following names
to instruct Landslide to assert the following corresponding predicates at the end of each test execution,
after all threads exit %
\footnote{Each predicate will be checked iff its first listed variable name is defined;
if that variable is defined, all others associated must also be;
any combination of the three first-listed variables
}.
(These cannot be implemented as asserts in the test code itself
without requiring the student to implement {\tt thr\_join()} and {\tt thr\_exit()}.)
\begin{itemize}
	\item {\tt magic\_global\_expected\_result == magic\_global\_value}
	\item {\tt magic\_expected\_sum == magic\_thread\_local\_value\_parent +} \\ {\tt magic\_thread\_local\_value\_child}
	\item {\tt magic\_expected\_sum\_minus\_result == magic\_thread\_local\_value\_parent +} \\ {\tt magic\_thread\_local\_value\_child - magic\_global\_value}
\end{itemize}

@@

\subsection{Test cases}

>>>>

\section{Pebbles}

@@

\subsection{Test cases}
\label{sec:education-pebbles-tests}

Landslide ships with several ``approved'' test cases,
i.e., programs copied from, derived from, or at least vaguely resembling
the tests distributed with P2,
which I curated to produce concurrent behaviour suitable for stateless model checking.
Some tests are crafted to target specific bugs which,
from personal experience as a TA, are common in many student submissions;
others are crafted to exercise generally concurrency-heavy code paths and uncover any number of unforeseen problems.
Many use some of the features/annotations described in \sect{\ref{sec:landslide-testcases}}.

@@

@@

\subsection{Test cases}
\label{sec:education-pintos-tests}

### 2018-06-15 thesis/landslide.tex
describe mysterious test case annos in landslide chapter

>>>>

\subsection{Test cases}
\label{sec:landslide-testcases}

Landslide depends on human intuition to construct a test case
that will produce both meaningful in quality and manageable in quantity thread interleavings.

The user may supply custom test cases
for Pebbles (under {\tt pebsim/p2-basecode})
by creating a file in {\tt 410user/progs} and adding it to config.mk as usual,
or for Pintos (under {\tt pebsim/pintos/src/tests/threads})
by creating a file and adding it to both {\tt tests.c} and {\tt Make.tests}.
Tests for the most common interactions during the P2 and Pintos projects are of course already supplied,
as described in \sect{\ref{sec:education-pebbles-tests}} and \sect{\ref{sec:education-pintos-tests}}.

Use of {\tt tell\_landslide()} annotations is not necessary,
although {\tt tell\_landslide\_\allowbreak{}preempt()} and {\tt tell\_landslide\_dump\_stack()}
may optionally be used at the user's convenience.
Additionally, the following ``secret'' annotations
are occasionally used in the pre-supplied test cases
to accomplish several mysterious goals described hereupon.

\subsubsection{Magic post-test assertions}

Test cases may define global variables of the following names
to instruct Landslide to assert the following corresponding predicates at the end of each test execution,
after all threads exit.
Each predicate will be checked iff its first listed variable name is defined;
if that variable is defined, all others associated must also be;
any combination of the three first-listed variables may be specified at the user's option.
\begin{itemize}
	\item {\tt magic\_global\_expected\_result == magic\_global\_value}
	\item {\tt magic\_expected\_sum == magic\_thread\_local\_value\_parent +} \\ {\tt magic\_thread\_local\_value\_child}
	\item {\tt magic\_expected\_sum\_minus\_result == magic\_thread\_local\_value\_parent +} \\ {\tt magic\_thread\_local\_value\_child - magic\_global\_value}
\end{itemize}
These could not be implemented as asserts in the test code itself
without requiring the student to implement {\tt thr\_join()} and {\tt thr\_exit()},
avoiding which is important for tests to be student-accessible earlier in the project implementation timeline.

\subsubsection{Misbehave}

Many of the P2 test cases can be seen to invoke the {\tt misbehave()} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff.
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

To illustrate, consider the following program.

% TODO
% TODO - make a progr that data races immediately after thr create, no mutex es or whatever.
% TODO - check if it rly happens under pure happens before? or if this is only a concern under limited HB.

Other {\tt misbehave} modes may be used, but are likely to have no effect,
since Landslide's thread-scheduling algorithm will override
any Pathos-internal scheduling priorities that may arise therefrom.
Hypothetically speaking, a reader with access to the top-secret Pebbles source code
could find further {\tt misbehave} documentation in its {\tt inc/misbehavior.h}.
